Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-21 talk/pepm2011/paper.tex
review the new section 4

The main insight to improve the code shown in the last section is that objects
in category 1 do not survive very long -- they are used only inside the loop and
there is no other outside reference to them. The optimizer identifies objects in
category 1 and removes the allocation of these objects, and all operations
manipulating them.

@@

To optimize the trace, it is traversed from beginning to end and an output trace
is produced at the same time. Every operation in the input trace is either
removed, or put into the output trace. Sometimes new operations need to be
produced as well. The optimizer can only remove operations that manipulate
objects that have been allocated within the trace, all others are copied to the
output trace untouched.

Looking at the example trace of Figure~\ref{fig:unopt-trace}, this is what
happens with the operations in lines 1-9. They are manipulating objects that
existed before the trace because they are passed in as arguments. Therefore the
optimizer just puts them into the output trace.

The following operations (lines 10-17) are more interesting:

@@

When the optimizer encounters a \lstinline{new}, it removes it optimistically,
and assumes that the object is in category 1. When the object escapes later, it
will be allocated at that point. The optimizer needs to keep track
of how the object that the operation creates looks like at various points in
the trace. This is done with the help of a \emph{static object}\footnote{Here
``static'' is meant in the sense of partial evaluation, \ie known at partial
evaluation time, not in the sense of ``static allocation'' or ``static
method''.}. The static object describes the shape of the object that would have
been allocated, \ie the type of the object and where the values that would be
stored in the fields of the allocated object come from.

@@

The \lstinline{guard_class} operations can be removed, since their argument is a
static object with the matching type \lstinline{BoxedInteger}. The
\lstinline{get} operations can be removed as well, because each of them reads a
field out of a static object. The results of the get operation are replaced with
what the static object stores in these fields: $i_{7}$ and $i_{8}$ are just
replaced by $i_{4}$ and -100. The only operation put into the optimized trace
is the addition:

>>>>

The main insight to improve the code shown in section \ref{sub:example} is that objects
in category 1 do not survive very long -- they are used only inside the loop and
there is no other outside reference to them. The optimizer identifies objects in
category 1 and removes the allocation of these objects, and all operations
manipulating them.

@@

To optimize the trace, it is traversed from beginning to end and an output trace
is produced at the same time. Every operation in the input trace is either
removed, or put into the output trace. Sometimes new operations need to be
produced as well. The optimizer can only remove operations that manipulate
objects that have been allocated within the trace, while all other operations are copied to the
output trace untouched.

Looking at the example trace of Figure~\ref{fig:unopt-trace}, the operations
in lines 1-9 are manipulating objects which existed before the trace and that
are passed in as arguments: therefore the optimizer just puts them into the
output trace.

The following operations (lines 10-17) are more interesting:
\anto{what about putting the line numbers also here? Moreover, figure 4 is put just after this listing, it could be confused}

@@

\anto{I don't like ``how the object looks like''. I don't know what else to say, though :-(} 
When the optimizer encounters a \lstinline{new}, it removes it optimistically,
and assumes that the object is in category 1. If later the optimizer finds that the object escapes, it
will be allocated at that point. The optimizer needs to keep track
of how the object that the operation creates looks like at various points in
the trace. This is done with the help of a \emph{static object}\footnote{Here
``static'' is meant in the sense of partial evaluation, \ie known at partial
evaluation time, not in the sense of ``static allocation'' or ``static
method''.}. The static object describes the shape of the object that would have
been allocated, \ie the type of the object and where the values that would be
stored in the fields of the allocated object come from.

@@

The \lstinline{guard_class} operations can be removed, since their argument is a
static object with the matching type \lstinline{BoxedInteger}. The
\lstinline{get} operations can be removed as well, because each of them reads a
field out of a static object. The results of the get operation are replaced with
what the static object stores in these fields: all the occurences of $i_{7}$ and $i_{8}$ in the trace are just
replaced by $i_{4}$ and -100. The only operation put into the optimized trace
is the addition:
